---
title: "Data Science HW-1"
author: "Kasturi Bhamidipati"
date: "09/23/2022"
output: github_document
---

First, we will load packages that are necessary to run this code! 

```{r, echo=FALSE}
library(tidyverse)
library(palmerpenguins)
library(skimr)
```

# Problem 1

### Summary of the dataset: 

```{r }
data("penguins", package = "palmerpenguins")
```

This is the Penguins data set: 

`r skim(penguins)`

- It has three factor variables, species, island and sex, and five numeric variables, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g and year. 

- Three species of penguins were studied in this dataset, Adelie, Gentoo and Chinstrap, across three     islands, Torgersen, Dream and Biscoe.

- Data was collected across the years 2007-2009.

- It has `r nrow(penguins)` observations and `r ncol(penguins)` variables, i.e., a total of 344 penguins were studied and data across 8 different aspects were collected for each penguin.

- The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm=TRUE),2)`mm. 

### Scatterplot: 

Here we have the scatter plot: 
  
  ```{r, warning=FALSE}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color= species))+geom_point()
ggsave("scatterplot_penguins.pdf")
```


# Problem 2

### Creating the dataframe 

Creating the dataframe as per the problem requirement: 

- x: a random sample of size 10 from a standard Normal distribution
- logic_x: indicates whether elements of the sample are greater than 0
- character_y: character vector of length 10
- factor_z: factor vector of length 10, with 3 different factor “levels”; high, medium and low


```{r}
problem2_df = 
  tibble(
     x = rnorm(n=10), 
    logic_x = x>0,
    character_y = c("1", "2", "3", "4", "5","6","7","8","9","10"),
    factor_z = factor(c("low", "high", "medium", "medium", "low", "low", "high", "medium", "high", "high"))
  )
```

### Taking the mean of each variable 

```{r}
mean(pull(problem2_df, x))
mean(pull(problem2_df, logic_x))
mean(pull(problem2_df, character_y))
mean(pull(problem2_df, factor_z))
```

When I try taking the means for my variables, I get warning messages for both the character variable and the factor variable. However, for the logical variable, since it is a Boolean operator that is indicating whether the elements in my sample are greater than zero, by using 0s and 1s (TRUE/FALSE), it is able to calculate the mean. 




